---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Ranystyle

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

Ranystyle is an R package designed to automate the extraction, parsing, and cleaning of bibliographic references from PDF and text documents. Utilizing the power of the 'anystyle' Ruby gem, it segments references and converts them into structured formats suitable for analysis and use.

You can cite this package as:
```{r}
citation("Ranystyle")
```


## Installation

You can install the development version of Ranystyle from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("agoutsmedt/Ranystyle")
```

## Example

Here's a basic example of how you might use Ranystyle to parse and clean references from a PDF document:

```{r example}
library(Ranystyle)
# Path to your PDF document
pdf_path <- system.file("extdata", package = "Ranystyle") 
files <- list.files(pdf_path)

# Extract references from the PDF
extracted_refs <- find_ref_to_df(input = paste0(pdf_path, "/", files[1]))

# Print the extracted references
print(extracted_refs)
```

## Credits

*anystyle* has been developed by [Alex Fenton](https://github.com/a-fent), [Sylvester Keil](https://github.com/inukshuk), [Johannes Krtek](https://github.com/flachware) and [Ilja Srna](https://github.com/namyra). *anystyle* is under copyright: Copyright 2011-2018 Sylvester Keil. All rights reserved. See the [Licence](https://github.com/inukshuk/anystyle-cli/blob/master/LICENSE) for details.

The logo of *Ranystyle* has been generated with DALL·E.
