---
title: "Using Ranystyle: Extracting, Parsing and Cleaning Bibliographic References"
author: "Aurélien Goutsmedt"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using Ranystyle: Extracting, Parsing and Cleaning Bibliographic References}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Ranystyle)
library(magrittr)
```

# Introduction

Ranystyle is an R package designed to automate the extraction, parsing, and cleaning of bibliographic references from PDF and text documents. It relies on [anystyle](https://anystyle.io) application.^[*anystyle* is under copyright: Copyright 2011-2018 Sylvester Keil. All rights reserved. See the [Licence](https://github.com/inukshuk/anystyle-cli/blob/master/LICENSE) for details.] This vignette provides a comprehensive guide to using Ranystyle's functionalities, including extracting references from PDFs, parsing references from text files, and cleaning and organizing reference data.

# Initial Setup: installing anystyle

Before using the functionalities to extract or parse references, you need to ensure that the [anystyle](https://anystyle.io) Ruby gem is installed on your system.

The `install_anystyle()` function simplifies the installation of the necessary 'anystyle' components. Ensure that the Ruby environment is properly configured and that gem can be called from the command line.
See more details on Ruby's installation [here](https://www.ruby-lang.org/fr/downloads/) and on
RubyGems [here](https://rubygems.org/pages/download?locale=fr).

```{r}
# Install anystyle
install_anystyle()
```

# Extracting References from PDFs and text files

The `find_ref()` function is used to extract references from PDF documents. It utilizes the 'anystyle' Ruby gem `find` function to analyze the document and identify reference sections from PDF. The `parse_ref()` function, coming from anystyle `parse` function, is designed to parse structured references from text files or from a vector of character. It reads the text, applies parsing rules, and converts the references into a structured format.

The package incorporates various academic PDFs with bibliographies, as well as the bibliographies of these PDFs extracted in a .txt format, in order to manipulate the different functions of the graph.

```{r}
# Path to the documents of the package
pdf_path <- system.file("extdata", package = "Ranystyle") 
files <- list.files(pdf_path)

files
```

You can save the parsed references in various format: `bib`, `json`, `xml` or `ref` (a text document with one reference per line). By setting `path = ""` in `find_ref()` or `parse_ref()`, you can directly extract the results in R, rather than saving the results. Let's see an example with a PDF.^[The `no_layout` parameter can be set to TRUE when the bibliography has several columns.]

```{r}
doc <- files[1]
print(doc)
extracted_refs <- find_ref(input = paste0(pdf_path, "/", doc),
                           path = "",
                           output_format = "bib",
                           no_layout = FALSE)

# Print the extracted references
cat(stringr::str_trunc(extracted_refs, 523))
```

Or we can parse reference from a .txt.

```{r}
doc <- files[2]
print(doc)
extracted_refs <- parse_ref(input = paste0(pdf_path, "/", doc),
                           path = "",
                           output_format = "xml")

# Print the extracted references
cat(stringr::str_trunc(extracted_refs, 615))
```

`parse_ref()` can work directly on a vector of character, with one reference per line.

```{r}
# a set of false references
references <- c(
  "Smith, J. (2020). Understanding Artificial Intelligence. AI Journal, 35(5), 123-145.",
  "Johnson, L., and Brown, C. (2018). Advances in Machine Learning. New York: Academic Press.",
  "Clark, E., & Wright, R. (2019). Robotics and Automation. London: Springer, 200-250.",
  "Davis, M. 2021. Quantum Computing: A New Era. Quantum Computing, 5(3):300-320.",
  "Adams, James, and Murphy, Finn. (2022). Exploring Virtual Reality. VR World, 10(7), 777-800."
)

extracted_refs <- parse_ref(input = references,
                           path = "",
                           output_format = "json")

# Print the extracted references
cat(stringr::str_trunc(extracted_refs, 640))
```

# From Reference to Data Frame

Depending on whether your start from PDFs or from .txt. or vectors, you can use two functions: `find_ref_to_df()` and `parse_ref_to_df()`. These functions create a data frame (a `tibble` more precisely) gathering the various information of the references extracted.

```{r}
doc <- files[1]
print(doc)
data_ref <- find_ref_to_df(input = paste0(pdf_path, "/", doc),
                           no_layout = FALSE)

print(data_ref)
```

`find_ref_to_df()` and `parse_ref_to_df()` can also take multiple PDF or .txt documents and create a data frame from them.

```{r}
pdfs <- files %>% 
  .[stringr::str_detect(., "\\.pdf$")]
data_ref <- find_ref_to_df(input = paste0(pdf_path, "/", pdfs),
                           clean_ref = FALSE,
                           no_layout = FALSE)

data_ref
```

As you may have seen, anystyle runs twice per document: the second time serve to extract the references in a .ref format, and is used to keep the whole reference in a column of the table. This could be useful for cleaning the data later. 

```{r}
data_ref %>% 
  dplyr::select(id_ref, full_ref)
```

The data extracted are still a bit messy. Most columns are in list format, with sometimes two or more element per reference (like two dates, or two titles). The `clean_ref()` function implements different cleaning processes: 

- It collapses non-essential columns (i.e. other than `date` and `title`), combining multiple entries into a single text string.
- It reorganizes `author` information, creating a full name from given, particle, and family components. Also add an `author_order` column. The column will remain in a list format.
- It separates and cleans dates and titles, organizing them into primary and additional information. The first date and first title are save in `date` and `title`. The other dates and titles are saved in `other_date` and `other_title`.
- It extracts the year from the date and places it in a separate column.
- It removes extraneous punctuation and trims whitespace from character columns.
- It relocates key columns to a more standardized order for easier analysis and readability.

```{r}
data_ref <- clean_ref(data_ref) 

data_ref
```

These cleaning steps are directly implemented in `find_ref_to_df()` and `parse_ref_to_df()`: you just need to set `clean_ref` to TRUE.


